Cross-Dataset Recommender System for \\Overcoming Selection Bias

### 2022-08-18 paper/paper.tex
Write related works

\section{Introduction}
Selection bias is one of the most prevalent sources of biases for recommender systems \cite{chen2020bias}. Selection bias happens when there is a pattern in the users' ratings that is unique to the training set. For example, in a recommender system for movies, users might mainly rate movies that are recommended to them, which is a small section of movies already tailored to the users' propensities. However, the environment the recommender system is deployed on contains all movies regardless of personal tastes. This discrepancy produces a misalignment between training and deployed settings, which is known as a distribution shift. Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms\cite{schnabel2016recommendations}.

When recommender systems are deployed in real-world platforms, it is arguably likely that the clients are able to collect rating-associated data from different source distributions. These data (or feedbacks) are either implicit or explicit \cite{aggarwal2016recommender}, while implicit data often comes in a highly-quantized, binary form and explicit data is usually less quantized. For example, a user's explicit rating of a movie is usually quantized to a number between 1 and 5, while the implicit feedback, such as if a user spontaneous searches for a movie, is often binary (more quantized.) We extend our hypothesis and argue that implicit data contains less or no selection bias compared to explicit data. This is because implicit data are often from users' spontaneous actions while explicit data are prejudiced toward the output of the recommender system in the previous feedback loop and users' innate biases.

@@

\section{Related Work}
Prior works on overcoming selection bias-induced distribution shift fall

>>>>

\section{Introduction}
Selection bias is one of the most prevalent sources of biases for recommender systems \cite{chen2020bias}. Selection bias happens when there is a pattern in the users' ratings that is unique to the training set. For example, in a recommender system for movies, users might mainly rate movies that are recommended to them, which is a small section of movies already tailored to the users' tastes \cite{pradel2012ranking}. However, the environment the recommender system is deployed on contains all movies regardless of personal tastes. This discrepancy produces a misalignment between training and deployed settings, which is known as a distribution shift. Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms\cite{schnabel2016recommendations}.

When recommender systems are deployed in real-world platforms, it is arguably likely that the clients are able to collect rating-associated data from different source distributions. These data (or feedbacks) are either implicit or explicit \cite{aggarwal2016recommender}, while implicit data often comes in a highly-quantized, binary form and explicit data is usually less quantized. For example, a user's explicit rating of a movie is usually quantized to a number between 1 and 5, while the implicit feedback, such as if a user spontaneous searches for a movie, is often binary (more quantized.) We extend our hypothesis and argue that implicit data contains less or no selection bias compared to explicit data. This is because implicit data are often from users' spontaneous actions while explicit data are prejudiced toward the output of the recommender system in the previous feedback loop and also users' innate biases.

@@

\section{Related Work}
Prior works on overcoming selection bias-induced distribution shift via a propensity-based approach begins with the seminal paper \cite{schnabel2016recommendations}, which introduces the IPS method into recommender systems. Follow-up works aim at providing a learning-based or behavioral model of user feedbacks for propensity estimation \cite{joachims2017unbiased} \cite{yang2018unbiased}, which remains the central concern of this approach. In this work we add to previous attempts the idea of using a more quantized, implicit dataset for more accurate propensity estimation.

The IPS method in our context is, essentially, a method of weighting training examples to correct the bias in the training data. Equivalent approaches such as importance weighting are widely used for domain adaptation in fields other than recommender systems \cite{sugiyama2007covariate} \cite{zhang2018importance}. Discussions on the IPS-based domain adaptation for countering selection bias, which is most relevant to recommender systems, remain limited.
